### 2016-12-16 intro.tex
no message

\heading{Formalizing abstract data structures.} Loosely speaking, we
formalize a data structure as a pair of algorithms $(\Rep,\Qry)$,
the representation and query-evaluation algorithms, respectively.
The randomized representation algorithm takes in an abstract
object~$S$ and produces public representation data $\pubaux$ and
private representation data $\privaux$.  Intuitively, $\pubaux$ is
public information, whereas $\privaux$ is meant to be known only to
the party running $\Qry$.  The deterministic query-evaluation
algorithm uses $\pubaux$ and $\privaux$ in order to (approximately)
evaluate a supported query $\qry$, which can in principle be an
arbitrary function.
%We say ``approximately'' evaluate permitted queries
%because $\pubaux$ and $\privaux$ may not contain all of the
%information sufficient to determine the true value of $\qry(S)$ for
%all~$S$.

@@

\heading{Case studies.} The example above offers a hint at the broad
range of abstract data structures our syntax supports.  In this
paper, we put the syntax and our security notions to work in several
case studies.  First, we analyze classical Bloom filters when the
hash functions are modeled as public random
oracles~\cite{broder2004network}, when the hash functions are
secretly-keyed PRFs~\cite{niedermeyer2014cryptanalysis,C:NaoYog15},
and when the hash functions are secretly keyed but fall short of
being PRFs, i.e., so-called ``linear-PRF'' hash
functions~\cite{kirsch2006less,niedermeyer2014cryptanalysis}.  Each
of these represents an in-use or proposed Bloom-filter variant. We
also consider garbled Bloom filters~\cite{CCS:DonCheWen13}, proposed
as a key piece of a private set-intersection protocol.

@@

\heading{Related work.}
\input{related-work}
%\jnote{See the discussion of related work in
%Naor-Yogev. We should cite [LN93]. I can't tell how relevant [MNS11]
%is. [HW13] seems relevant, though I don't think it is captured by
%our current definitions.}

>>>>

\heading{Formalizing abstract data structures. } Loosely speaking,
we formalize a data structure as a pair of algorithms $(\Rep,\Qry)$,
the representation and query-evaluation algorithms, respectively.
The randomized representation algorithm takes in an abstract
object~$S$ and produces public representation data $\pubaux$ and
private representation data $\privaux$.  Intuitively, $\pubaux$ is
public information, whereas $\privaux$ is meant to be known only to
the party running $\Qry$.  The deterministic query-evaluation
algorithm uses $\pubaux$ and $\privaux$ in order to (approximately)
evaluate a supported query $\qry$, which can in principle be an
arbitrary function.
%We say ``approximately'' evaluate permitted queries
%because $\pubaux$ and $\privaux$ may not contain all of the
%information sufficient to determine the true value of $\qry(S)$ for
%all~$S$.

@@

\heading{Case studies. } The example above offers a hint at the
broad range of abstract data structures our syntax supports.  In
this paper, we put the syntax and our security notions to work in
several case studies.  First, we analyze classical Bloom filters
when the hash functions are modeled as public random
oracles~\cite{broder2004network}, when the hash functions are
secretly-keyed PRFs~\cite{niedermeyer2014cryptanalysis,C:NaoYog15},
and when the hash functions are secretly keyed but fall short of
being PRFs, i.e., so-called ``linear-PRF'' hash
functions~\cite{kirsch2006less,niedermeyer2014cryptanalysis}.  Each
of these represents an in-use or proposed Bloom-filter variant. We
also consider garbled Bloom filters~\cite{CCS:DonCheWen13}, proposed
as a key piece of a private set-intersection protocol.

@@

\heading{Related work. }
\input{related-work}
%\jnote{See the discussion of related work in
%Naor-Yogev. We should cite [LN93]. I can't tell how relevant [MNS11]
%is. [HW13] seems relevant, though I don't think it is captured by
%our current definitions.}

### 2016-12-16 results.tex
no message

%%%%%%%%%%%%%%%%% LINEAR-PRF %%%%%%%%%%%%%%%%%%%%
\begin{figure}[tp]
\centering
\hfpagess{.5}{.5}
{
\algorithmv{$\Rep(S)$}\\
$M \gets 0^m$; $K \getsr \calK$\\
for $x \in S$\\
%\jnote{Fix}\tsnote{Fix what?}\\
\nudge $f_1 \gets F_{K}(\langle 1,x \rangle)$, $f_2 \gets F_{K}(\langle 2,x \rangle)$\\
\nudge for $j \in \{1,2,\ldots,k\}$\\
\nudge\nudge $h_j \gets f_1 + j\cdot f_2 \bmod m$\\
\nudge\nudge $M[h_j] \gets 1$\\
$\pubaux \gets M$, $\privaux \gets \langle K \rangle$\\
Return $(\pubaux,\privaux)$\\

\medskip
\algorithmv{$\Qry(\pubaux,\privaux)$}\\
$M\gets\pubaux$, $K \gets \privaux$\\
%\jnote{Fix}\tsnote{Fix what?} \\
$f_1 \gets F_{K}(\langle 1,x \rangle)$, $f_2 \gets F_{K}(\langle 2,x \rangle)$\\
for $j \in \{1,2,\ldots,k\}$\\
\nudge $h_j \gets f_1 + j\cdot f_2 \bmod m$\\
\nudge if $M[h_j] \neq 1$ then Return 0\\
Return 1
}
{
\algorithmv{$\Rep(S)$}\\
$\pubaux \gets \emptystring$\\
$K_1,K_2, \getsr \calK$\\
for each $x \in S$\\
\nudge $M_x \gets 0^m$\\
\nudge $B_x \gets \mathsf{bigram}(x)$\\
\nudge for each $b \in B_x$\\
\nudge\nudge $f_1 \gets F_{K}(\langle 1,b \rangle)$, $f_2 \gets F_{K}(\langle 2,b \rangle)$\\
\nudge\nudge for $j \in [k]$\\
\nudge\nudge\nudge $i \gets f_1+j\cdot f_2 \bmod m$\\
\nudge\nudge\nudge $M_x[i] \gets 1$\\
\nudge $\pubaux \gets \langle\pubaux, M_x \rangle$\\
$\privaux \gets \langle K_1,K_2 \rangle$\\
Return $(\pubaux,\privaux)$\\

\medskip
\algorithmv{$\Qry(\pubaux,\privaux,q_x)$}\\
$B \gets \mathsf{bigram}(x)$\\
$\overline{M} \gets 0^m$\\
for each $b \in B$\\
\nudge for $j \in [k]$\\
\nudge\nudge $f_1 \gets F_{K}(\langle 1,b \rangle)$, $f_2 \gets F_{K}(\langle 2,b \rangle)$\\
\nudge\nudge $i \gets f_1+j\cdot f_2 \bmod m$\\
\nudge\nudge $\overline{M}[i] \gets 1$\\
if $\overline{M} \in \pubaux$ then Return 1 \\
Return 0
}
\caption{{\bf Left:} $\Rep$ and $\Qry$ algorithms for the linear-PRF
  data structure $\setprim_{\mathrm{lin}}$. {\bf Right:}
Data structure $\setprim_{\bilin}$ from Niedermeyer et
al.~\cite{niedermeyer2014cryptanalysis}. The $\mathsf{bigram}$
algorithm takes a string~$x=x_1,x_2,\ldots,x_\ell$ as input, and
returns the set of bigrams for the left- and right-padded string
$\sqcup x \sqcup$, i.e. $\{\sqcup x_1,
x_1x_2,\ldots,x_{\ell-1}x_\ell,x_\ell\sqcup\}$. Both schemes use a
function family $F\colon \calK \times \bits^* \to [m]$.
}\label{fig:lin-and-bi-lin} \label{fig:niedermeyer}
\end{figure}

@@

%JK
%%%%%%%%%%%%%%%%% BIGRAM LINEAR-PRF (AKA "NIEDERMEYER")  %%%%%%%%%%%%%%%%%%%%
\subsection{Bigram Linear-PRF Construction}
%\heading{Bigram Linear-PRF construction (``Niedermeyer''). }
%\heading{Correctness.}
%\todo{Correctness theorem for this scheme.  Should follow pretty closely the one for ``plain'' linear-PRF.}\tsnote{Actually not sure this is that interesting, but for parallel structure one would want to give it.}
Consider the set-membership data structure given in
Figure~\ref{fig:lin-and-bi-lin} (right side), which we call the
\emph{bigram linear-PRF} construction.  This construction is of
practical interest, as it has been suggested as a way to address the
privacy-preserving record-linkage problem, in particular for linking
common records in medical
databases~\cite{niedermeyer2014cryptanalysis}. %,schnell2011novel}.
\jnote{We should really give more discussion, explaining the goal of
``privacy-preserving record linkage'' and describing the protocol
(at a high level).}

\heading{Privacy. } We only address the privacy of this
construction, as the correctness follows immediately from the
correctness of the Linear-PRF construction.  The privacy, however,
does not.  Here we exhibit an efficient SS-privacy attack even when
simulator is provided nearly the entire set.  Consider the following
``all-but-last'' leakage function. On input a
set~$S=\{x_1,x_2,\ldots,x_n\}$, it assigns $x'_n \gets
x_n[1:|x_n|-1]$ and returns
$\mathsf{leak_{abl}}(S)=\{x_1,x_2,\ldots,x'_n\}$.

>>>>

%%%%%%%%%%%%%%%%% LINEAR-PRF %%%%%%%%%%%%%%%%%%%%
\begin{figure}[tp]
\centering
\hfpagess{.45}{.45}
{
\algorithmv{$\Rep(S)$}\\
$M \gets 0^m$; $K \getsr \calK$\\
for $x \in S$\\
%\jnote{Fix}\tsnote{Fix what?}\\
\nudge $f_1 \gets F_{K}(\langle 1,x \rangle)$, $f_2 \gets F_{K}(\langle 2,x \rangle)$\\
\nudge for $j \in \{1,2,\ldots,k\}$\\
\nudge\nudge $h_j \gets f_1 + j\cdot f_2 \bmod m$\\
\nudge\nudge $M[h_j] \gets 1$\\
$\pubaux \gets M$, $\privaux \gets \langle K \rangle$\\
Return $(\pubaux,\privaux)$\\

\medskip
\algorithmv{$\Qry(\pubaux,\privaux, q_x)$}\\
$M\gets\pubaux$, $K \gets \privaux$\\
$f_1 \gets F_{K}(\langle 1,x \rangle)$, $f_2 \gets F_{K}(\langle 2,x \rangle)$\\
for $j \in \{1,2,\ldots,k\}$\\
\nudge $h_j \gets f_1 + j\cdot f_2 \bmod m$\\
\nudge if $M[h_j] \neq 1$ then Return 0\\
Return 1
}
{
\algorithmv{$\Rep(S)$}\\
$\pubaux \gets \emptystring$; $K \getsr \calK$\\
for each $x \in S$\\
\nudge $M_x \gets 0^m$\\
\nudge $B_x \gets \mathsf{bigram}(x)$\\
\nudge for each $b \in B_x$\\
\nudge\nudge $f_1 \gets F_{K}(\langle 1,b \rangle)$, $f_2 \gets F_{K}(\langle 2,b \rangle)$\\
\nudge\nudge for $j \in \{1, 2, \ldots, k\}$\\
\nudge\nudge\nudge $h_j \gets f_1+j\cdot f_2 \bmod m$\\
\nudge\nudge\nudge $M_x[h_j] \gets 1$\\
\nudge $\pubaux \gets \langle\pubaux, M_x \rangle$\\
$\privaux \gets \langle K \rangle$\\
Return $(\pubaux,\privaux)$\\

\medskip
\algorithmv{$\Qry(\pubaux,\privaux,q_x)$}\\
$B \gets \mathsf{bigram}(x)$\\
$M_x \gets 0^m$\\
for each $b \in B$\\
\nudge $f_1 \gets F_{K}(\langle 1,b \rangle)$, $f_2 \gets F_{K}(\langle 2,b \rangle)$\\
\nudge for $j \in \{1, 2, \ldots, k\}$\\
\nudge\nudge $h_j \gets f_1+j\cdot f_2 \bmod m$\\
\nudge\nudge $M_x[h_j] \gets 1$\\
if $M_x \in \pubaux$ then Return 1 \\
Return 0
}
\caption{{\bf Left:} $\Rep$ and $\Qry$ algorithms for the linear-PRF
  data structure $\setprim_{\mathrm{lin}}$. {\bf Right:}
Data structure $\setprim_{\bilin}$ from Niedermeyer et
al.~\cite{niedermeyer2014cryptanalysis}. \jnote{Can we not describe
this in terms of $\setprim_{\mathrm{lin}}$?} The $\mathsf{bigram}$
algorithm takes a string~$x=x_1,x_2,\ldots,x_\ell$ as input, and
returns the set of bigrams for the left-and-right-padded string
$\sqcup x \sqcup$, i.e. $\{\sqcup x_1,
x_1x_2,\ldots,x_{\ell-1}x_\ell,x_\ell\sqcup\}$. Both schemes use a
function family $F\colon \calK \times \bits^* \to [m]$.
}\label{fig:lin-and-bi-lin} \label{fig:niedermeyer}
\end{figure}

@@

%%%%%%%%%%%%%%%%% BIGRAM LINEAR-PRF (AKA "NIEDERMEYER")  %%%%%%%%%%%%%%%%%%%%
\subsection{Bigram Linear-PRF Construction}
%\heading{Bigram Linear-PRF construction (``Niedermeyer''). }
%\heading{Correctness.}
%\todo{Correctness theorem for this scheme.  Should follow pretty closely the one for ``plain'' linear-PRF.}\tsnote{Actually not sure this is that interesting, but for parallel structure one would want to give it.}
Consider the set-membership data structure given in
Figure~\ref{fig:lin-and-bi-lin} (right side), which we call the
\emph{bigram linear-PRF} construction. \jnote{Really a bad
name\ldots} This construction is of interest as it has been
suggested as a way to address the privacy-preserving record-linkage
problem, in particular for linking common records in medical
databases~\cite{niedermeyer2014cryptanalysis}. %,schnell2011novel}.
\jnote{We should really give more discussion, explaining the data
structure in words, and also discussing the goal of
``privacy-preserving record linkage'' and describing the protocol
for doing so (at a high level).}

\heading{Privacy. } We only address the privacy of this
construction, as the correctness follows immediately from the
correctness of the linear PRF-hash construction. \jnote{Well, with
some loss of parameters.} The privacy, however, does not.
%JK
Here we
exhibit an efficient SS-privacy attack even when simulator is
provided nearly the entire set. Consider the following
``all-but-last'' leakage function. On input a
set~$S=\{x_1,x_2,\ldots,x_n\}$, it assigns $x'_n \gets
x_n[1:|x_n|-1]$ and returns
$\mathsf{leak_{abl}}(S)=\{x_1,x_2,\ldots,x'_n\}$.
